# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

We read in the data and reformat the "date" column to be R date objects.

```{r}
setwd("~/GitHub/RepData_PeerAssessment1")
df <- read.csv("activity.csv")
df$date <- as.Date(df$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

First we clean the data set by removing NA values, then we find the total number of steps per day.

```{r}
library(reshape2)
df2 <- df[!is.na(df$steps),]
totSteps <- tapply(df2$steps,df2$date,sum)
totSteps <- totSteps[!is.na(totSteps)]
```

Here is a histogram of the total number of steps per day.

```{r}
hist(totSteps)
```

The mean number of steps taken per day is:

```{r}
mean(totSteps)
```

The median number of steps taken per day is:

```{r}
median(totSteps)
```

## What is the average daily activity pattern?

```{r}
meanSteps <- tapply(df2$steps,df2$interval,mean)
plot(names(meanSteps),meanSteps,type="l",xlab="5-minute interval",ylab="Number of Steps Taken")
```

To find the time interval with the highest average number of steps, we use the which.max function:

```{r}
which.max(meanSteps)
```

## Imputing missing values

For missing (NA) values, we will first compute a "default day" of data, consisting of the mean for 
the missing time interval.

```{r}
defaultDay <- data.frame(mean=meanSteps)
```

Now we create a new data set that is equal to the old one, but with the NA values filled in with these defaults.

```{r}
df3 <- df
df3$steps[is.na(df3$steps)] <- defaultDay[as.character(df3$interval[is.na(df3$steps)]),]
```

Let's compare the total number of steps per day of this new data set to the original.

```{r}
totSteps3 <- tapply(df3$steps,df3$date,sum)
par(mfrow = c(1,2))
hist(totSteps,main="Original Data Set",xlab="Total Steps per Day")
hist(totSteps3,main="New Data Set",xlab="Total Steps per Day")
```

There are slight differences in the new data set, but the overall pattern seems to be the same.

## Are there differences in activity patterns between weekdays and weekends?

First we separate weekends from weekdays.

```{r}
df3$weekendQ <- (weekdays(df3$date) == "Saturday") | (weekdays(df3$date) == "Sunday")
```

Now we reshape the data and compute means.

```{r}
meltedData <- melt(df3,id=c("weekendQ","interval"),measure.vars=c("steps"))
meanData <- dcast(meltedData,weekendQ + interval ~ variable,mean)
meanData$dayType <- factor(meanData$weekendQ,labels=c("Weekday","Weekend"))
```

Finally, we create a lattice plot of our weekday and weekend data sets.

```{r}
library(lattice)
xyplot(steps ~ interval | dayType, data=meanData,type="l",layout=c(1,2),xlab="Interval",ylab="Number of Steps")
```

